---
title: "GSE142512 DNA methylation data"
author: "Paul Ruiz Pinto"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{GSE142512 DNA methylation data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document
---

# GSE142512 DNA methylation data

`GSE142512` provides access to processed DNA methylation data derived from GEO
accession
[GSE142512](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE142512).
The source data include GPL13534 Illumina HumanMethylation450 and GPL21145
Infinium MethylationEPIC arrays.

The ExperimentHub resources are `RangedSummarizedExperiment` objects with beta,
M, and CN assays. The 450K resource contains 485,512 CpG loci across 184
samples from 84 subjects. The EPIC resource contains 865,859 CpG loci across
211 samples from 90 subjects. Genomic ranges use hg19 coordinates.

The source publication is:

Johnson RK, Vanderlinden LA, Dong F, Carry PM, Seifert J, Waugh K, Shorrosh H,
Fingerlin T, Frohnert BI, Yang IV, Kechris K, Rewers M, Norris JM.
Longitudinal DNA methylation differences precede type 1 diabetes. Scientific
Reports. 2020;10:3721. doi:10.1038/s41598-020-60758-0.

# Accessing the resources

The data can be located in `ExperimentHub` using the GEO accession or the
package name. The loaded objects can then be inspected with standard
`SummarizedExperiment` accessors.

```{r access-resource, eval=FALSE}
library(ExperimentHub)
library(GSE142512)
library(SummarizedExperiment)

eh <- ExperimentHub()
query(eh, "GSE142512")

se450k <- GSE142512(platform = "450K")
seEPIC <- GSE142512(platform = "EPIC")

se450k <- GSE142512_GPL13534_450K()
seEPIC <- GSE142512_GPL21145_EPIC()

dim(se450k)
assayNames(se450k)
colData(se450k)
rowRanges(se450k)
metadata(se450k)
```

The package metadata describe their sources.

```{r local-metadata}
metadataFile <- system.file("extdata", "metadata.csv", package = "GSE142512")
metadata <- utils::read.csv(metadataFile, stringsAsFactors = FALSE)
metadata[, c("Title", "Genome", "SourceType", "RDataClass", "DispatchClass")]
```

# Data availability

The serialized `RangedSummarizedExperiment` objects are hosted on Zenodo:

```{r zenodo-location}
paste0(metadata$Location_Prefix, metadata$RDataPath)
```

# Session information

```{r sessionInfo}
sessionInfo()
```
